% SHARD-ID: CA-04-HILBERT-SPACE-COMPILER-CONTRACT
% SHARD-TITLE: A Hilbert-Space Compiler Contract
% SHARD-SUMMARY: States the mini north star for the pre-work sequence: expressions in the many-particle grammar should compile to Hilbert carriers, observable-algebra policies, symmetries, and evidence metadata.
% SHARD-SUMMARY: Separates syntax from semantics and records the typed output slots needed before the grammar can support sectors, quotients, and later category-valued data.
% SHARD-SUMMARY: Treats the regular-expression analogy as a proposal while keeping every constructor responsible for explicit Hilbert-space and observable-algebra data.
% SHARD-KEYWORDS: Hilbert-space compiler, grammar, observable algebra, regular expressions, syntax, semantics, evidence metadata

\section{A Hilbert-Space Compiler Contract}
\label{sec:hilbert-space-compiler-contract}

\subsection{Status and Local Anchors}

\textbf{Status: Proposal / contract shard.}  This shard introduces no new
mathematical classification theorem.  It records a design target for the
pre-work sequence around CA-02 and CA-03: a typed grammar of many-particle
expressions should be compilable to Hilbert-space data and observable-algebra
data with explicit evidence status.

Local anchors are CA-02 for the finite AND/OR/vacuum grammar, CA-03 for the
Hilbert direct-sum Fock completion, and \texttt{CONVENTIONS.md} entries (f) and
(g) for symmetry and sector semantics.  The first executable check for this
sequence is the finite averaging projector in \texttt{test/runtests.jl}, backed
by \texttt{src/CftAnyons.jl}.

\subsection{Mini North Star}

Proposal: build a \emph{Hilbert-space compiler}.  The user supplies an expression
in a typed grammar, together with the atom table and required policies; the
compiler returns the Hilbert carrier, its observable algebra policy, its symmetry
representation data, and any sector projection or quotient presentation that was
requested.

The motivating analogy is ``regular expressions for quantum mechanics'': a small
syntax has compositional semantics.  The analogy is only a prompt.  Here the
syntax is typed and semantics-bearing; every constructor must say what it does to
Hilbert carriers, particle gradings, observable algebras, symmetries, and
evidence metadata.

\subsection{Input Record}

A compiled expression is not just a string.  It has an input record:
\[
  \mathsf{Input}
  =
  (\mathsf{Atoms},\ E,\ \mathsf{Completion},\ \mathsf{ObsPolicy},\
   \mathsf{Symmetry},\ \mathsf{SectorPolicy}).
\]

\begin{description}
\item[\(\mathsf{Atoms}\).]
A table of atoms.  At minimum, each atom names a Hilbert space \(K_x\).  Later
entries may include a group representation, a particle degree, a source locator,
or category data.
\item[\(E\).]
A grammar expression built from \(0,\one,K_x,\operatorname{AND},\operatorname{OR}\),
Fock, and later sector constructors.
\item[\(\mathsf{Completion}\).]
Whether \(\oplus,\otimes\) are finite Hilbert operations, Hilbert completions
\(\hoplus,\hotimes\), or a finite cutoff.  CA-03 fixes the first infinite Fock
completion used here.
\item[\(\mathsf{ObsPolicy}\).]
The observable-algebra choice.  There is no silent default: full
\(\BH(H)\), block-preserving, tensor-local, symmetry-invariant, and compressed
sector algebras are different outputs.
\item[\(\mathsf{Symmetry}\).]
Optional unitary or projective-unitary representation data, handled according to
CONVENTIONS.md (f).
\item[\(\mathsf{SectorPolicy}\).]
Optional projection, subspace, or quotient data; see CONVENTIONS.md (g).
\end{description}

\subsection{Output Record}

The intended output is a typed record, not only a Hilbert-space dimension:
\[
  \mathsf{Compile}(E)
  =
  (H_E,\ \{H_{E,n}\}_{n\ge0},\ \Alg_E,\ U_E,\ P_E,\ Q_E,\ \mathsf{Evidence}_E).
\]

\begin{description}
\item[\(H_E\).]
The Hilbert carrier.  CA-02 and CA-03 define the first carrier semantics.
\item[\(\{H_{E,n}\}\).]
The particle-number grading when the atom table declares particle degrees.
\item[\(\Alg_E\).]
The selected observable algebra.  This slot is mandatory once any observable
claim is made.
\item[\(U_E\).]
The propagated symmetry representation, if symmetry data were supplied.
\item[\(P_E\).]
The selected sector projection, if a sector constructor was used.
\item[\(Q_E\).]
The quotient presentation, if one was requested and its closed relation subspace
was named.
\item[\(\mathsf{Evidence}_E\).]
Source-local, checked, proposal, question, or rejected evidence tags for the
pieces above.
\end{description}

\subsection{Observable-Algebra Policies}

The carrier alone does not determine the observable algebra relevant to a model.
For example, \(H\oplus K\) supports at least:
\[
  \BH(H\oplus K),\qquad
  \BH(H)\oplus \BH(K),\qquad
  \text{off-diagonal creation/annihilation blocks between }H\text{ and }K.
\]
All three are legitimate in different contexts; none is forced by the word
``OR'' alone.  Similarly, \(H\otimes K\) supports the full algebra
\(\BH(H\otimes K)\) and the tensor-local subalgebra generated by
\(\BH(H)\otimes I\) and \(I\otimes\BH(K)\).

Compiler rule: no report statement may infer an observable algebra merely from a
carrier constructor.  The statement must name the observable policy, or remain a
carrier-only statement.

\subsection{Lamport Dependency Skeleton}

{\small
\[
\setlength{\arraycolsep}{2pt}
\begin{array}{rcl}
D10 &:& \mathsf{Input}=(\mathsf{Atoms},E,\mathsf{Completion},\mathsf{ObsPolicy},
        \mathsf{Symmetry},\mathsf{SectorPolicy}).\\
D11 &:& \mathsf{Compile}(E)=(H_E,\{H_{E,n}\},\Alg_E,U_E,P_E,Q_E,
        \mathsf{Evidence}_E).\\
D12 &:& \mathsf{ObsPolicy}\in\{\text{carrier-only},\BH(H),\text{block},
        \text{tensor-local},\text{invariant},\text{compressed}\}.\\
T7 &:& \text{A carrier claim needs }H_E;\ \text{an observable claim also needs }\Alg_E.\\
T8 &:& \text{The compiler analogy is a proposal, not a theorem of canonical quantization.}
\end{array}
\]
}

\subsection{Boundary}

This contract does not yet compile fusion-category labels, anyonic braid
statistics, Hamiltonians, continuum limits, or CFT data.  It only fixes the
pre-work target: future shards should be able to say exactly which input record
they compile and which output slots they have actually justified.
