% SHARD-ID: CA-05-SYMMETRY-DECORATED-GRAMMAR
% SHARD-TITLE: Symmetry-Decorated Many-Particle Grammar
% SHARD-SUMMARY: Adds ordinary and projective-unitary symmetry data to atoms and propagates it through OR, AND, and full Fock constructions.
% SHARD-SUMMARY: Uses central extensions as the pre-work representation of projective symmetries, avoiding unsourced cocycle arithmetic in the grammar.
% SHARD-SUMMARY: Records the compatibility requirements needed before multiple atomic Hilbert spaces can share a single compiled symmetry action.
% SHARD-KEYWORDS: symmetry, unitary representation, projective representation, central extension, tensor product, direct sum, Fock

\section{Symmetry-Decorated Many-Particle Grammar}
\label{sec:symmetry-decorated-grammar}

\subsection{Sources and Dependencies}

This shard extends CA-02 and CA-03 by decorating atoms with symmetry data.  It
uses CONVENTIONS.md (f).

% Source: references/cft/Schottenloher2008/Schottenloher2008.md:1524
%   unitary representation = continuous homomorphism G -> U(H).
% Source: references/cft/Schottenloher2008/Schottenloher2008.md:1528
%   projective representation = continuous homomorphism G -> U(P).
% Source: references/cft/Schottenloher2008/Schottenloher2008.md:1570
%   a projective representation determines a central extension E of G and a lift.
% Source: references/cft/Schottenloher2008/Schottenloher2008.md:1618
%   in the topological setting, the lift is a continuous unitary representation.
% Source: references/text/PenneysUnitaryFusionCategories.md:323
%   Rep^\dagger(G) has Hilbert spaces with unitary G-actions and equivariant maps.

\subsection{Decorated Atoms}

Fix a topological group \(G\).  A symmetry-decorated atom is a pair
\[
  (K_x,U_x),\qquad U_x:G\to U(K_x),
\]
where \(K_x\) is the atomic Hilbert space and \(U_x\) is a unitary
representation.  If the intended symmetry is projective, the pre-work convention
is to choose a central extension
\[
  1\to U(1)\to E\to G\to 1
\]
and use an honest unitary representation \(U_x:E\to U(K_x)\).  The expression is
then compiled over \(E\), with a note that the induced physical symmetry is
projective over \(G\).

The vacuum atom is \((\mathbb C,\mathbf 1)\), where \(\mathbf 1(g)z=z\).  The
zero atom carries the unique representation on \(\zero\).

\subsection{Propagation Through OR and AND}

If \(E,F\) have compiled carriers and representations \((H_E,U_E)\) and
\((H_F,U_F)\) for the same group or chosen central extension, define
\[
  H_{\operatorname{OR}(E,F)}=H_E\oplus H_F,\qquad
  U_{\operatorname{OR}(E,F)}(g)=U_E(g)\oplus U_F(g),
\]
and
\[
  H_{\operatorname{AND}(E,F)}=H_E\hotimes H_F,\qquad
  U_{\operatorname{AND}(E,F)}(g)=U_E(g)\hotimes U_F(g).
\]
For finite-dimensional factors, \(\hotimes=\otimes\).  In infinite-dimensional
Hilbert spaces, the completed tensor product is the one already required by
CA-03.

These formulas are the Hilbert-space version of the monoidal structure on
\(\Rep^\dagger(G)\): direct sums are block actions and tensor products are
diagonal actions.  If two atoms use different groups, different central
extensions, or incompatible projective multipliers, the compiler must stop and
ask for a common symmetry object before forming a single \(U_E\).

\subsection{Propagation Through Full Fock}

For an ordinary unitary representation \(U:G\to U(H)\), define the full-Fock
action
\[
  \Fock(U)(g)
  =
  \bigoplus_{n\ge0} U(g)^{\hotimes n}
  \quad\text{on}\quad
  \Fock(H)=\hoplus_{n\ge0}H^{\hotimes n},
\]
with \(U(g)^{\hotimes0}=\operatorname{id}_{\mathbb C}\).  This preserves the
particle-number grading.  In this pre-work shard, no Shale--Stinespring-type
implementability theorem is claimed for general canonical field algebras; the
formula is only the sectorwise tensor-power representation on the full Fock
carrier.  Implementing a one-particle unitary on a chosen CAR/CCR observable
algebra is a separate field-algebra question.

\subsection{Intertwiners and Symmetry-Preserving Compilation}

A morphism between decorated atoms or compiled expressions is symmetry-preserving
only when it is an intertwiner:
\[
  T U_E(g)=U_F(g)T\quad\text{for all }g.
\]
This matters for compiler optimisations.  Rebracketing and distributivity
isomorphisms used by CA-02 are acceptable in the decorated grammar only when they
also intertwine the propagated actions.  The standard Hilbert-space associators,
unitors, symmetries, and distributivity maps do so for diagonal direct-sum and
tensor-product actions.

\subsection{Lamport Dependency Skeleton}

{\small
\[
\setlength{\arraycolsep}{2pt}
\begin{array}{rcl}
D13 &:& \text{decorated atom }(K_x,U_x),\ U_x:G\to U(K_x).\\
D14 &:& \text{projective symmetry over }G\text{ is compiled via }U_x:E\to U(K_x)
        \text{ for a named central extension }E.\\
D15 &:& U_{E\operatorname{OR}F}=U_E\oplus U_F,\quad
        U_{E\operatorname{AND}F}=U_E\hotimes U_F.\\
D16 &:& \Fock(U)=\hoplus_{n\ge0}U^{\hotimes n},\quad U^{\hotimes0}=1_{\mathbb C}.\\
T9 &:& \text{The propagated symmetry preserves the CA-02/CA-03 particle grading.}\\
T10 &:& \text{A single compiled }U_E\text{ requires a common group or central extension.}
\end{array}
\]
}

\subsection{Boundary}

This shard does not yet impose invariance, select charge sectors, form quotient
spaces, or define symmetry-invariant observable algebras.  Those are sector
operations, not mere symmetry propagation, and are treated in CA-06.
