% SHARD-ID: CA-07-EXCHANGE-STATISTICS-LAYER
% SHARD-TITLE: Exchange Symmetry and Statistics as Sector Selection
% SHARD-SUMMARY: Applies the sector machinery to the permutation action on tensor slots, with symmetric and antisymmetric sectors arising from chosen irreps of S_n.
% SHARD-SUMMARY: Keeps internal symmetries separate from exchange symmetry and records how the two actions coexist on n-particle sectors.
% SHARD-SUMMARY: Defers para-statistics and genuinely anyonic braid statistics until their source, convention, and projector data are fixed.
% SHARD-KEYWORDS: exchange symmetry, statistics, symmetric Fock, antisymmetric Fock, permutation group, sector projector, anyonic deferral

\section{Exchange Symmetry and Statistics as Sector Selection}
\label{sec:exchange-statistics-layer}

\subsection{Sources and Dependencies}

This shard discharges part of the deferral in CA-03: ordinary Bose/Fermi
statistics are treated as choices of irreducible exchange-symmetry sectors on
the full distinguishable Fock carrier.  It depends on CA-05 for internal
symmetry propagation and CA-06 for projection/subspace/quotient semantics.

% Source: literature/md/1910.10619/1910.10619.md:39
%   Fock presentations often mix exchange symmetry and indefinite particle number.
% Source: literature/md/1910.10619/1910.10619.md:41
%   Bose-/Fermi-Fock space can be obtained by imposing exchange equivalence.
% Source: literature/md/2010.11121/2010.11121.md:258
%   bosonic Fock space uses symmetric tensor powers.
% Source: references/cft/Schottenloher2008/Schottenloher2008.md:1642
%   fermionic Fock/spinor space is built from exterior powers.
% Source: references/cft/Schottenloher2008/Schottenloher2008.md:6672
%   symmetric powers can also be presented as a quotient of tensor algebra.
% Source: references/text/TrebstShortIntroductionFibonacciAnyons.txt:42
%   two-dimensional anyon exchange can be Abelian phases or non-Abelian matrices.

\subsection{Two Symmetry Layers}

There are two distinct actions on many-particle carriers:
\[
  \text{internal symmetry }G,\qquad
  \text{exchange symmetry }S_n.
\]
If \(U:G\to U(H)\) is an internal one-particle representation, then on the
\(n\)-particle distinguishable sector \(H^{\hotimes n}\) it acts diagonally:
\[
  U_n(g)=U(g)^{\hotimes n}.
\]
The permutation group \(S_n\) acts by permuting tensor slots:
\[
  \Pi_n(\sigma)(v_1\otimes\cdots\otimes v_n)
  =
  v_{\sigma^{-1}(1)}\otimes\cdots\otimes v_{\sigma^{-1}(n)}.
\]
The two actions commute:
\[
  \Pi_n(\sigma)U_n(g)=U_n(g)\Pi_n(\sigma).
\]
Thus internal charge sectors and exchange-statistics sectors can be imposed in
either order in the ordinary tensor-slot setting.

\subsection{Irrep-Labeled Exchange Projectors}

The representation \(\Pi_n:S_n\to U(H^{\hotimes n})\) is the primary datum.
Sector projections are supplied by irreducible representation data of \(S_n\),
not by specialised grammar constructors.  For a one-dimensional unitary
character \(\chi:S_n\to U(1)\), CA-06 applied to the twisted representation
\(\overline{\chi}\,\Pi_n\) gives
\[
  P_{\chi,n}
  =
  \frac1{n!}\sum_{\sigma\in S_n}\overline{\chi(\sigma)}\,\Pi_n(\sigma),
\]
whose range is the \(\chi\)-sector
\[
  \{w\in H^{\hotimes n}:\Pi_n(\sigma)w=\chi(\sigma)w\ \forall\sigma\in S_n\}.
\]
For the trivial character \(1\), this is the symmetric projector:
\[
  P_{\mathrm s,n}
  =
  \frac1{n!}\sum_{\sigma\in S_n}\Pi_n(\sigma),
\]
and for the sign character \(\operatorname{sgn}\), this is the antisymmetric
projector:
\[
  P_{\mathrm a,n}
  =
  \frac1{n!}\sum_{\sigma\in S_n}\operatorname{sgn}(\sigma)\Pi_n(\sigma).
\]
Their ranges are the symmetric and antisymmetric tensor powers:
\[
  \Ran(P_{\mathrm s,n})=H^{\hotimes_s n},\qquad
  \Ran(P_{\mathrm a,n})=\bigwedge^n H.
\]
The corresponding Fock sectors are
\[
  \Gamma_{\mathrm s}(H)=\hoplus_{n\ge0}\Ran(P_{\mathrm s,n}),\qquad
  \Gamma_{\mathrm a}(H)=\hoplus_{n\ge0}\Ran(P_{\mathrm a,n}).
\]
Thus ``Bose'' and ``Fermi'' name the trivial- and sign-irrep choices for ordinary
exchange symmetry.  The compiler should record the selected irrep/character and
the resulting projection, not a special Bose/Fermi syntax node.

\subsection{Quotient Presentations}

For a one-dimensional character \(\chi\), the sector can also be presented as a
quotient of the full tensor algebra by the closed linear span of relations
\[
  \Pi_n(\sigma)v-\chi(\sigma)v.
\]
The trivial character recovers symmetric relations; the sign character recovers
antisymmetric signed relations.  In the compiler, these quotient presentations
are secondary: the primary sector object is the orthogonal projector, because it
fixes the inherited Hilbert structure and the compressed observable algebra.

\subsection{Para and Anyonic Boundaries}

Higher-dimensional irreducible \(S_n\)-sectors require central idempotents or
matrix-block projectors for the group algebra of \(S_n\).  This is exactly the
same representation/irrep principle, but this shard records it as a proposal
rather than a completed construction; a later shard must source the convention
and projector normalization before using it.

Genuinely anyonic exchange is not an \(S_n\)-projection problem on ordinary tensor
slots.  It involves braid-group or categorical data and, in this project, must be
handled together with the fusion-category conventions for associators, braiding,
and \(F/R\)-symbols.  No braid-statistics projector is defined here.

\subsection{Lamport Dependency Skeleton}

{\small
\[
\setlength{\arraycolsep}{2pt}
\begin{array}{rcl}
D21 &:& U_n(g)=U(g)^{\hotimes n}\quad\text{(internal symmetry on }H^{\hotimes n}).\\
D22 &:& \Pi_n:S_n\to U(H^{\hotimes n})\quad\text{permutes tensor slots}.\\
D23 &:& P_{\chi,n}=n!^{-1}\sum_{\sigma}\overline{\chi(\sigma)}\Pi_n(\sigma)
        \quad(\chi\text{ a one-dimensional }S_n\text{ character}).\\
L3 &:& \Pi_n(\sigma)U_n(g)=U_n(g)\Pi_n(\sigma).\\
L4 &:& P_{\chi,n}\text{ is an orthogonal projection by CA-06 finite averaging.}\\
T13 &:& \Gamma_{\mathrm s},\Gamma_{\mathrm a}\text{ are the trivial/sign irrep sectors inside full Fock.}\\
T14 &:& \text{Anyonic exchange remains deferred to categorical braid data.}
\end{array}
\]
}

\subsection{Compiler Consequence}

The compiler now has its first nontrivial sector family: given the exchange
representation \(\Pi_n:S_n\to U(H^{\hotimes n})\), a selected irrep character
\(\chi\) supplies \(P_{\chi,n}\), its range, its quotient relations, and the
chosen compressed observable algebra.  There is no primitive
\(\operatorname{Bose}\), \(\operatorname{Fermi}\), or \(\operatorname{Fock}\)
grammar node in this account; Fock is the derived graded representation of CA-05,
and Bose/Fermi are names for particular irrep choices of exchange symmetry.
