% SHARD-ID: CA-11-CONTINUUM-BULK-SYMMETRY-TARGET
% SHARD-TITLE: The Continuum Bulk Symmetry Target
% SHARD-SUMMARY: Fixes the continuum symmetry target as Poincare or Euclidean motion symmetry with projective-unitary implementation.
% SHARD-SUMMARY: Derives the Poincare vector-field brackets from the metric-preserving vector fields rather than importing an unsourced table.
% SHARD-SUMMARY: Pins the boost-time-translation-to-momentum relation up to the sign convention used by the lattice candidate layer.
% SHARD-KEYWORDS: Poincare algebra, Lorentz algebra, Euclidean group, boost, momentum, vector fields, Lie bracket

\section{The Continuum Bulk Symmetry Target}
\label{sec:continuum-bulk-symmetry-target}

\subsection{Status, Sources, and Dependencies}

\textbf{Status: sourced definitions plus checked local derivation.}  The
Poincare group and the projective-unitary implementation are sourced.  The
commutator table below is a local vector-field derivation from the sourced
metric-preserving description, and its signs are pinned by a Julia check.

Dependencies: CA-05, CA-10, and CONVENTIONS.md (h), (k).

% Source: references/cft/Schottenloher2008/Schottenloher2008.md:578 -- affine conformal Killing fields.
% Source: references/cft/Schottenloher2008/Schottenloher2008.md:599 -- o(p,q) condition omega^T g + g omega = 0.
% Source: references/cft/Schottenloher2008/Schottenloher2008.md:790 -- in 1+1, the Poincare group is R semidirect R^2.
% Source: references/cft/Schottenloher2008/Schottenloher2008.md:1297 -- Poincare as SO(1,3) semidirect R^4.
% Source: references/cft/Schottenloher2008/Schottenloher2008.md:4092 -- Poincare group generated by Lorentz and translations.
% Source: references/cft/Schottenloher2008/Schottenloher2008.md:4102 -- projective Poincare representation lifted to a unitary covering representation.
% Source: references/cft/Schottenloher2008/Schottenloher2008.md:4040--4050 -- Stone convention U(t)=exp(-itA).
% Source: references/cft/Schottenloher2008/Schottenloher2008.md:4113--4117 -- translation subgroup generators and P_0=H.
% Convention: CONVENTIONS.md (k) -- map from vector-field brackets to self-adjoint commutator signs.
% Checked: test/runtests.jl, "Poincare vector-field brackets".

\subsection{Group-Level Target}

For a \(D=d+1\) dimensional Lorentzian target, Schottenloher describes the
Poincare group as the semidirect product of the Lorentz group and translations:
\[
  P = SO_0(1,d)\ltimes \mathbb R^{d+1},
\]
with the usual replacement by the relevant covering group when the quantum
symmetry is implemented projectively.  The translation subgroup is abelian and
has commuting self-adjoint generators \(P_\mu\).  In the source convention,
\[
  P_0=H,\qquad P_a\text{ are spatial momenta }(a=1,\ldots,d).
\]

For Euclidean two-dimensional CFT language, the Euclidean motion group is
generated by rotations and translations.  This gives the Wick-rotated analogue:
rotations and translations are the geometric target before scale/conformal
extensions are added.

\subsection{Vector-Field Derivation}

Let \(\eta=\operatorname{diag}(1,-1,\ldots,-1)\), and lower indices by
\(x_\mu=\eta_{\mu\nu}x^\nu\).  Use vector fields
\[
  P_\mu=\partial_\mu,\qquad
  M_{\mu\nu}=x_\mu\partial_\nu-x_\nu\partial_\mu,\qquad M_{\mu\nu}=-M_{\nu\mu}.
\]
The source gives the metric-preserving Lie algebra condition
\(\omega^T\eta+\eta\omega=0\).  The \(M_{\mu\nu}\) form the corresponding
coordinate vector fields, while \(P_\mu\) are translations.

Compute the Lie bracket of vector fields.  Since the \(P_\mu\) have constant
coefficients,
\[
  [P_\mu,P_\nu]=0.
\]
For the mixed bracket,
\[
\begin{aligned}
  [M_{\mu\nu},P_\rho]
  &= -\partial_\rho(x_\mu)\partial_\nu
     +\partial_\rho(x_\nu)\partial_\mu\\
  &= \eta_{\nu\rho}P_\mu-\eta_{\mu\rho}P_\nu.
\end{aligned}
\]
For Lorentz-Lorentz brackets,
\[
\begin{aligned}
  [M_{\mu\nu},M_{\rho\sigma}]
  &=
    \eta_{\nu\rho}M_{\mu\sigma}
    -\eta_{\mu\rho}M_{\nu\sigma}\\
  &\quad
    +\eta_{\mu\sigma}M_{\nu\rho}
    -\eta_{\nu\sigma}M_{\mu\rho}.
\end{aligned}
\]
The Julia check evaluates representative brackets using this convention.  For
example in \(3+1\) dimensions,
\[
  [M_{01},P_0]=-P_1,\qquad [M_{01},P_1]=-P_0.
\]

\subsection{Boost-Time Relation}

With \(K_a:=M_{0a}\), the vector-field convention gives
\[
  [K_a,P_0]=-P_a.
\]
Equivalently, \([P_0,K_a]=P_a\).  The passage from this vector-field statement to
self-adjoint commutator notation is the separate sign convention in
CONVENTIONS.md (k): Stone generators use \(U_X(t)=\exp(-itA_X)\), the spacetime
flow is implemented passively, and the resulting target is
\[
  i[H,K_a]=P_a,\qquad i[P_b,K_a]=\delta_{ab}H.
\]
Spatial rotations carry the additional convention \(J_{ab}=-A_{M_{ab}}\), which
is why the Gaussian one-particle table has \(i[K_a,K_b]=J_{ab}\) even though the
active vector fields obey \([M_{0a},M_{0b}]=-M_{ab}\).  CA-12 keeps finite-lattice
signs explicit under CONVENTIONS.md (h).

\subsection{Boundary}

This shard does not claim that a finite lattice has an exact Poincare
representation.  It only identifies the continuum algebraic target whose
relations a sequence of lattice candidates should approximate after scaling.
