% SHARD-ID: CA-13-POSITION-DEPENDENT-BULK-GENERATORS
% SHARD-TITLE: Position-Dependent Bulk Generators
% SHARD-SUMMARY: Generalizes the one-dimensional first-moment idea to scalar ramps, vector fields, translations, rotations, and boosts on lattices with positions.
% SHARD-SUMMARY: Distinguishes sourced continuum geometry from proposal-level lattice formulas for higher-dimensional bulk generators.
% SHARD-SUMMARY: Explains why rotations need momentum-density candidates, while boosts can start directly from the energy density.
% SHARD-KEYWORDS: position-dependent translation, lattice rotation, lattice boost, vector field, momentum density, higher dimensions

\section{Position-Dependent Bulk Generators}
\label{sec:position-dependent-bulk-generators}

\subsection{Status, Sources, and Dependencies}

\textbf{Status: proposal shard with local derivation pattern.}  The continuum
geometric statements are sourced.  The higher-dimensional lattice formulas are
compiler targets, not established convergence claims.

Dependencies: CA-10--CA-12 and CONVENTIONS.md (h).

% Source: references/cft/Schottenloher2008/Schottenloher2008.md:582 -- constant vector fields flow to translations.
% Source: references/cft/Schottenloher2008/Schottenloher2008.md:695 -- Euclidean Killing field X(z)=c+i theta z.
% Source: references/cft/Schottenloher2008/Schottenloher2008.md:697 -- Euclidean motions z -> c + exp(i theta)z.
% Source: references/cft/Schottenloher2008/Schottenloher2008.md:4788 -- Euclidean motions generated by rotations.
% Source: references/cft/Schottenloher2008/Schottenloher2008.md:4794 -- Euclidean motions generated by translations.
% Checked: CA-12 and test/runtests.jl for the scalar-ramp nearest-neighbour identity.

\subsection{Continuum Picture}

Translations are flows of constant vector fields.  In the Euclidean plane, the
infinitesimal motion
\[
  X(z)=c+i\theta z
\]
contains both translations (\(c\)) and rotations (\(i\theta z\)).  Written in
real coordinates, the rotation part is the position-dependent vector field
\[
  (x,y)\mapsto (-\theta y,\theta x).
\]
This is the precise continuum content of the slogan:
\[
  \text{rotation}=\text{position-dependent translation}.
\]

\subsection{Scalar Ramps}

Suppose lattice density terms are attached to positions \(r\in\Lambda\subset
\mathbb R^d\).  For a scalar function \(f:\Lambda\to\mathbb R\), define
\[
  K_f:=\sum_{r\in\Lambda} f(r)h_r.
\]
The same derivation as CA-12 gives, for commuting non-overlapping densities,
\[
  i[H,K_f]
  =
  \sum_{\{r,s\}}(f(s)-f(r))\,i[h_r,h_s],
\]
where the sum is over unordered overlapping pairs after a choice of orientation.
This is a discrete-gradient response.  Linear ramps
\[
  f_a(r)=r^a
\]
are therefore the first candidates for spatial momentum components.

\subsection{Boost Candidates}

Boosts can be attempted directly from energy density:
\[
  K_a^{\mathrm{lat}}:=\sum_{r\in\Lambda} r^a h_r.
\]
Then \(i[H,K_a^{\mathrm{lat}}]\) gives the \(a\)-direction current candidate.
This is the higher-dimensional version of the one-dimensional derivation, but
it depends on the chosen density supports and orientations.

\subsection{Rotation Candidates}

Rotations require one more layer.  In continuum notation,
\[
  J_{ab}=\int (x_a p_b(x)-x_b p_a(x))\,dx.
\]
Thus a lattice rotation candidate should be built from local momentum-density or
current candidates, not from \(h_r\) alone.  A provisional compiler rule is:
\[
  J_{ab}^{\mathrm{lat}}
  \approx
  \sum_r \bigl(r_a p_b^{\mathrm{lat}}(r)-r_b p_a^{\mathrm{lat}}(r)\bigr),
\]
where \(p_a^{\mathrm{lat}}\) is itself obtained from scalar ramps and commutators
as above.

\subsection{Compiler Moral}

The lattice Hilbert-space compiler cannot infer these operators from the carrier
alone.  It needs a lattice geometry, a local Hamiltonian-density decomposition,
and a target vector-field algebra.  The output is not an exact symmetry by
default; it is a candidate list with proof obligations.

\subsection{Boundary}

No cell-orientation convention is fixed here.  No periodic wrapping convention
is fixed here.  No higher-dimensional convergence theorem is fixed here.  Those
choices must be made in a later shard before any numerical or categorical
lattice model uses these candidates as checked symmetries.
