% SHARD-ID: CA-14-KOO-SALEUR-PROTOTYPE
% SHARD-TITLE: Koo-Saleur as the Prototype
% SHARD-SUMMARY: Uses the original Koo-Saleur source to record the lattice Virasoro prototype from Hamiltonian-density modes.
% SHARD-SUMMARY: Records the Koo-Saleur pattern: Fourier modes of h_j plus commutator corrections approximate Virasoro generators.
% SHARD-SUMMARY: Separates the original conjectural/numerical support from the later rigorous free-fermion/OAR convergence theorem.
% SHARD-KEYWORDS: Koo-Saleur, Virasoro, Hamiltonian density, Fourier modes, commutator correction, OAR

\section{Koo--Saleur as the Prototype}
\label{sec:koo-saleur-prototype}

\subsection{Status, Sources, and Dependencies}

\textbf{Status: source-local prototype plus later rigorous precedent.}  The
original Koo--Saleur paper is registered here and supports the lattice Virasoro
prototype as a conjectural construction with numerical and analytic checks for
XXZ/RSOS/Temperley--Lieb settings.  This shard keeps that separate from the
local lattice-fermion/OAR treatment, which states and proves a free-fermion
Koo--Saleur convergence result.

Dependencies: CA-10--CA-13.

% Source: references/lattice-symmetry/KooSaleur1993/source/9312156.tex:36--45 -- original scope: XXZ/RSOS scaling limits, conjectural stress tensor, mostly numerical checks.
% Source: references/lattice-symmetry/KooSaleur1993/source/9312156.tex:606--648 -- Hamiltonian-limit density and ground-state subtraction e_infty.
% Source: references/lattice-symmetry/KooSaleur1993/source/9312156.tex:704--712 -- T_xy component involves nearest-neighbour commutators.
% Source: references/lattice-symmetry/KooSaleur1993/source/9312156.tex:877--925 -- weighted Hamiltonians, Fourier expansion, vacuum subtraction, and normal ordering.
% Source: references/lattice-symmetry/KooSaleur1993/source/9312156.tex:949--976 -- single Fourier components map to L_n +/- \bar L_{-n} in the double scaling limit.
% Source: references/lattice-symmetry/KooSaleur1993/source/9312156.tex:1030--1062 -- commutator calculation, central term, and finite-L non-closure.
% Source: references/lattice-symmetry/KooSaleur1993/source/9312156.tex:1068--1108 -- double-limit process and commutator/scaling-limit caveat.
% Source: references/lattice-symmetry/KooSaleur1993/source/9312156.tex:1623--1673 -- vertex-model lattice Virasoro generators l_n and \bar l_{-n}.
% Source: references/lattice-symmetry/KooSaleur1993/source/9312156.tex:2219--2266 -- RSOS conjecture with the same lattice quantities.
% Source: references/lattice-symmetry/KooSaleur1993/source/9312156.tex:2380--2396 -- conclusion: numerical evidence, exact Bethe-ansatz work still needed, conjectured homomorphism.
% Source: references/text/CFTFromLatticeFermions.txt:385--388 -- KS convergence relies on correct scaling representation and vacuum subtraction.
% Source: references/text/CFTFromLatticeFermions.txt:395--404 -- transverse-field Ising even subalgebra, KS convergence, and extra odd/full-covariance work.
% Source: references/text/CFTFromLatticeFermions.txt:2128--2133 -- continuum H is integral of Hamiltonian density.
% Source: references/text/CFTFromLatticeFermions.txt:2133--2138 -- h is local and given by chiral and anti-chiral stress tensors.
% Source: references/text/CFTFromLatticeFermions.txt:2163--2173 -- Fourier modes H_k are linear combinations of Virasoro generators.
% Source: references/text/CFTFromLatticeFermions.txt:2182--2195 -- Koo-Saleur proposal introduces lattice Fourier modes of Hamiltonian density.
% Source: references/text/CFTFromLatticeFermions.txt:2197--2217 -- KS approximants and their smeared versions.
% Source: references/text/CFTFromLatticeFermions.txt:2233--2235 -- smeared KS approximants converge to smeared Virasoro generators.
% Source: references/text/CFTFromLatticeFermions.txt:2247--2254 -- unitary exponentials and automorphic dynamics from smeared approximants.
% Source: references/text/CFTFromLatticeFermions.txt:2256--2257 -- those converge to localized diffeomorphism implementers/actions.

\subsection{Continuum Side}

In a two-dimensional CFT on the circle, the Hamiltonian is expressed as an
integral of a Hamiltonian density,
\[
  H=\int h(x)\,dx.
\]
The source identifies \(h(x)\) as a local quantum field built from the chiral and
anti-chiral energy-momentum tensors \(T,\bar T\).  Its Fourier modes \(H_k\)
are linear combinations of the Virasoro generators \(L_k,\bar L_k\).

Thus in this special setting the desired spacetime symmetries are not abstract:
they are generated by modes of the stress tensor.

\subsection{Lattice Side}

The Koo--Saleur move is to imitate the continuum construction using the lattice
Hamiltonian density.  From a lattice density
\[
  h^{(N)}:\Lambda_N\to \mathcal A_N
\]
one forms Fourier modes \(H_k^{(N)}\).  In the original Temperley--Lieb
presentation, the Hamiltonian-limit density is represented by
ground-state-subtracted \(e_j-e_\infty\) terms, while the second stress-tensor
component is represented by nearest-neighbour commutators
\([e_j,e_{j+1}]\).  The lattice Virasoro candidates \(l_n,\bar l_{-n}\) are
therefore Fourier sums of
\[
  (e_j-e_\infty)\ \pm\ \text{constant}\cdot [e_j,e_{j+1}],
\]
with the \(c/24\) zero-mode shift.  The later free-fermion/OAR Definition 4.1
uses the same structural pattern: density modes plus commutator corrections
against the lattice Hamiltonian mode \(H_0^{(N)}\) produce lattice analogues of
\(L_k\) and \(\bar L_k\).

The important structural point for our programme is:
\[
  \text{lattice symmetry generator}
  =
  \text{weighted Hamiltonian density}
  +
  \text{commutator correction}.
\]
This is the same motif as CA-12, but now with Fourier weights and a Virasoro
target rather than a first moment and a Poincare translation target.

\subsection{Scaling Representation Matters}

The original source already builds in ground-state subtraction, normal ordering,
and a double-limit process; it also warns that finite-\(L\) lattice quantities
do not close as the Virasoro algebra and that a scaling limit of a commutator is
not generally the commutator of scaling limits.  The later rigorous
free-fermion/OAR source similarly makes convergence depend on the correct
scaling-limit representation and on subtracting ground-state expectation values,
corresponding to normal ordering.  Therefore the compiler must not output only
formal sums.  It must also output:
\[
\begin{array}{ll}
  &\text{state sequence},\\
  &\text{renormalisation/scaling maps},\\
  &\text{vacuum subtraction policy},\\
  &\text{operator topology target}.
\end{array}
\]

\subsection{What This Foreshadows}

Koo--Saleur is a prototype for the lattice-symmetry compiler:
\[
  \{h_j\} \longmapsto
  \{H_k^{(N)},L_k^{(N)},\bar L_k^{(N)}\}
  \longrightarrow
  \{L_k,\bar L_k\}.
\]
The first arrow is algebraic and finite-scale.  The second is analytic and
requires a scaling theorem.

\subsection{Boundary}

The original Koo--Saleur source supports the prototype as a conjectural and
numerically checked construction for XXZ/RSOS/Temperley--Lieb lattice models; it
does not prove a general rigorous convergence theorem for arbitrary microscopic
models.  The convergence theorem cited here is still the later
free-fermion/OAR result, and arbitrary anyon chains remain outside both sourced
claims.
