% SHARD-ID: CA-15-LATTICE-SYMMETRY-ACCEPTANCE-TESTS
% SHARD-TITLE: Acceptance Tests for Lattice Symmetry Candidates
% SHARD-SUMMARY: Turns lattice symmetry guesses into proof obligations: locality, self-adjointness, commutator closure, scaling convergence, and observable covariance.
% SHARD-SUMMARY: Uses OAR and continuum-limit sources to stress that translations can be recovered while Lorentz/conformal covariance is extra work.
% SHARD-SUMMARY: Defines the evidence payload a compiler must carry before calling a lattice candidate a continuum symmetry.
% SHARD-KEYWORDS: acceptance tests, scaling limit, Lieb-Robinson, covariance, Poincare, Lorentz, compiler evidence

\section{Acceptance Tests for Lattice Symmetry Candidates}
\label{sec:lattice-symmetry-acceptance-tests}

\subsection{Status, Sources, and Dependencies}

\textbf{Status: specification shard.}  This shard does not prove convergence for
new models.  It states the checks required before a candidate generator is
promoted from proposal to checked continuum symmetry data.

Dependencies: CA-10--CA-14.

% Source: references/text/OARWavelets.txt:10 -- rigorous OAR scheme constructs free field with continuous translations.
% Source: references/text/OARWavelets.txt:13 -- causality follows from Lieb-Robinson bounds.
% Source: references/text/OARWavelets.txt:389 -- translations/dynamics/locality/Lieb-Robinson section.
% Source: references/text/OARWavelets.txt:393 -- continuous extension of spatial translations from dyadic translations.
% Source: references/text/OARWavelets.txt:398 -- translation generators are usual momentum operators.
% Source: references/text/OARWavelets.txt:405 -- Lorentz/conformal convergence requires further work.
% Source: references/text/OsborneContinuumLimitsQuantumLatticeSystems.txt:328 -- Poincare invariance is not automatic.
% Source: literature/md/2010.11121/2010.11121.md:241 -- time-zero net axioms.
% Source: literature/md/2010.11121/2010.11121.md:243 -- Poincare covariance is an additional requirement.

\subsection{Local Finite-Scale Checks}

At fixed lattice size \(N\), a candidate generator \(G_N\) must first pass
finite operator checks:
\[
  G_N=G_N^*,\qquad
  G_N\text{ is built from named local density terms},\qquad
  \mathrm{supp}(G_N)\text{ policy is explicit}.
\]
For unbounded thermodynamic-limit candidates, the finite check is only the
first layer; domains must later be named.

\subsection{Algebraic Checks}

Given a target Lie algebra with generators \(X_a\), the lattice candidates
\(G_a^{(N)}\) should satisfy commutator closure in a controlled sense:
\[
  i[G_a^{(N)},G_b^{(N)}]
  \approx
  f_{ab}^{\ \ c}G_c^{(N)}+\text{central or boundary terms}.
\]
The approximation relation must be specified: exact finite-size equality,
matrix-element convergence, strong resolvent convergence, strong operator
convergence on a core, or convergence after compression to a low-energy sector.

\subsection{Scaling Checks}

The OAR source shows that continuous spatial translations can be reconstructed
from discrete dyadic translations in a suitable scaling limit, with the usual
momentum operators as generators.  It also states that analogous convergence
for Lorentz or conformal transformations requires further work.  This is the
main warning for this programme:
\[
  \text{translation recovery}\not\Rightarrow
  \text{Poincare recovery}.
\]

Therefore every candidate boost or rotation needs its own scaling statement.
For a sequence \(G_N\), the report must say:
\[
  G_N \xrightarrow{\ ?\ } G
\]
and name the topology, state sequence, renormalisation maps, and dense domain or
core on which the statement is intended.

\subsection{Observable Covariance Checks}

The target is not only an operator on states.  It is a symmetry action on the
observable system.  The evidence payload should include
\[
  \alpha_N^G(A)=e^{itG_N}Ae^{-itG_N}
\]
or the infinitesimal commutator \(i[G_N,A]\), together with the expected
transformation of local algebras.  For translations this means moving supports.
For rotations it means rotating supports.  For boosts it means mixing time and
space in the continuum representation.

\subsection{Acceptance Payload}

A lattice generator candidate is accepted into the compiler only with:
\[
\begin{array}{ll}
1.&\text{density provenance: the exact }h_j\text{ and supports},\\
2.&\text{finite algebra checks: self-adjointness and commutators},\\
3.&\text{renormalisation maps and state sequence},\\
4.&\text{operator topology or matrix-element convergence target},\\
5.&\text{observable covariance statement},\\
6.&\text{boundary/central/vacuum-subtraction policy}.
\end{array}
\]

\subsection{Boundary}

Time-zero net axioms and Lieb-Robinson bounds give important locality control,
but local causality is not the same as full Poincare covariance.  The compiler
must keep these evidence levels separate.
