% SHARD-ID: CA-16-LATTICE-SYMMETRY-COMPILER-INTERFACE
% SHARD-TITLE: A Compiler Interface for Lattice Symmetry Candidates
% SHARD-SUMMARY: Extends the Hilbert-space compiler contract with lattice geometry, Hamiltonian densities, candidate generators, and symmetry evidence.
% SHARD-SUMMARY: Specifies successful and failed outputs for boosts, translations, rotations, and Koo-Saleur-style modes.
% SHARD-SUMMARY: Keeps candidate generation separate from proof of continuum covariance.
% SHARD-KEYWORDS: lattice compiler, symmetry compiler, Hamiltonian density, generator candidates, evidence payload, compile error

\section{A Compiler Interface for Lattice Symmetry Candidates}
\label{sec:lattice-symmetry-compiler-interface}

\subsection{Status, Sources, and Dependencies}

\textbf{Status: compiler specification shard.}  This shard extends CA-08 at the
level of output slots.  It is not yet an implementation.

Dependencies: CA-08 and CA-10--CA-15.

% Source: references/text/CFTFromLatticeFermions.txt:95 -- local terms plus observable algebra are needed for scaling-limit meaning.
% Source: references/text/CFTFromLatticeFermions.txt:2182--2195 -- lattice Hamiltonian-density Fourier modes.
% Source: references/text/CFTFromLatticeFermions.txt:2197--2217 -- KS approximants from density modes and commutators.
% Source: references/text/OARWavelets.txt:393--398 -- discrete translations extend to arbitrary translations in the continuum limit.
% Source: references/text/OARWavelets.txt:405--406 -- Lorentz/conformal convergence remains further work.

\subsection{Input Contract}

The lattice symmetry compiler receives a compiled Hilbert/QM system plus:
\[
\begin{array}{ll}
\Lambda_N &: \text{finite lattice, graph, or cell complex},\\
x &: \Lambda_N\to\mathbb R^d\text{ position map},\\
\mathcal A_N &: \text{observable algebra with locality map},\\
H_N=\sum_r h_r &: \text{self-adjoint local Hamiltonian density},\\
T &: \text{target symmetry algebra or group},\\
S_N &: \text{state/sector sequence and scaling maps, if known}.
\end{array}
\]
If the Hamiltonian is supplied only as a black-box matrix, the compiler must
fail for boost/rotation generation: the density decomposition is part of the
mathematical data.

\subsection{Candidate Generation}

For one-dimensional open chains, the first candidate layer is deterministic:
\[
  K_x=\sum_j x_jh_j,\qquad
  P_x=i[H,K_x].
\]
For uniform nearest-neighbour density this reduces to the CA-12 current sum.

For higher-dimensional lattices, scalar ramps \(f_a(r)=r^a\) generate candidate
momentum components
\[
  P_a^{\mathrm{lat}}:=i[H,K_{f_a}].
\]
Rotation candidates are then built only after the momentum-density/current layer
has its own convention and evidence:
\[
  J_{ab}^{\mathrm{lat}}
  \approx
  \sum_r(r_a p_b^{\mathrm{lat}}(r)-r_b p_a^{\mathrm{lat}}(r)).
\]
Until that layer is fixed, this is a deferred proposal slot rather than a
formula emitted by the compiler.
Koo--Saleur-style modes use Fourier weights instead of polynomial weights and
include the commutator corrections specified by the cited Definition 4.1.

\subsection{Output Contract}

The compiler output for each candidate generator must include:
\[
\begin{array}{ll}
\text{name} &: \text{for example }P_a,K_a,J_{ab},L_k,\bar L_k,\\
\text{formula} &: \text{finite expression in }h_r\text{ and commutators},\\
\text{domain policy} &: \text{finite matrix, core, or unbounded deferral},\\
\text{locality support} &: \text{site/bond/cell support accounting},\\
\text{status} &: \text{Proposal, Source-local, or Checked},\\
\text{evidence} &: \text{source lines, derivation, test, or run artifact},\\
\text{continuum target} &: \text{target generator and expected bracket}.
\end{array}
\]

\subsection{Failure Modes}

The compiler must fail loudly when:
\[
\begin{array}{ll}
1.&H\text{ has no density decomposition},\\
2.&\Lambda_N\text{ has no position map but a boost/rotation is requested},\\
3.&\text{periodic boundary positions are requested without a branch convention},\\
4.&\text{the target algebra requires a source/convention not yet fixed},\\
5.&\text{a convergence claim is requested without scaling maps or states}.
\end{array}
\]

\subsection{Compiler Principle}

The finite expression and the continuum theorem are different compilation
products.  The first can often be generated algebraically from
\((\Lambda_N,x,\{h_r\})\).  The second is a proof obligation involving scaling
maps, states, and operator topology.  The compiler should output both slots, but
it must never fill the proof slot with intuition.
