% SHARD-ID: CA-19-GALILEAN-VECTOR-FIELD-ALGEBRA
% SHARD-TITLE: The Galilei Algebra as Newtonian Vector Fields
% SHARD-SUMMARY: Derives the unextended Galilei Lie algebra from vector fields on Newtonian spacetime with absolute time.
% SHARD-SUMMARY: Fixes the sign convention for time translations, spatial translations, boosts, and rotations, with representative brackets checked in Julia.
% SHARD-SUMMARY: Separates the unextended vector-field algebra from the projective mass-central extension treated in CA-20.
% SHARD-KEYWORDS: Galilei algebra, Newtonian spacetime, vector fields, boosts, rotations, Lie bracket, checked derivation

\section{The Galilei Algebra as Newtonian Vector Fields}
\label{sec:galilean-vector-field-algebra}

\subsection{Status, Sources, and Dependencies}

\textbf{Status: checked local derivation.}  The source anchor is that the
Galilei group is the nonrelativistic free-particle symmetry group.  The bracket
table below is derived here from vector fields and checked by the Julia test
suite.

Dependencies: CA-18 and CONVENTIONS.md (i).

% Source: references/cft/Schottenloher2008/Schottenloher2008.md:1443 -- Galilei group for free particles in classical nonrelativistic mechanics.
% Checked: test/runtests.jl, "Galilei vector-field brackets".

\subsection{Generators}

Use Newtonian spacetime coordinates
\[
  (t,x_1,\ldots,x_d).
\]
The convention fixed in CONVENTIONS.md (i) is:
\[
\begin{array}{rcll}
  H   &=& \partial_t, &\text{time translation},\\
  P_a &=& \partial_a, &\text{spatial translation},\\
  G_a &=& t\,\partial_a, &\text{Galilean boost},\\
  J_{ab} &=& x_a\partial_b-x_b\partial_a,\quad a<b,
       &\text{spatial rotation}.
\end{array}
\]
Here \(a,b,c\in\{1,\ldots,d\}\).  The boost vector field says that a boost
changes position by a velocity times time while leaving absolute time fixed.

\subsection{Translation and Boost Brackets}

The translation fields have constant coefficients, hence
\[
  [H,P_a]=0,\qquad [P_a,P_b]=0.
\]
Since \(G_a=t\partial_a\),
\[
\begin{aligned}
  [H,G_a]
    &= [\partial_t,t\partial_a]\\
    &= \partial_t(t)\partial_a\\
    &= P_a.
\end{aligned}
\]
The reverse bracket is therefore \([G_a,H]=-P_a\).  Spatial translations do not
differentiate the coefficient \(t\), so
\[
  [P_a,G_b]=0,\qquad [G_a,G_b]=0.
\]
This is the unextended vector-field algebra.  It is not yet the projective
quantum algebra; CA-20 adds the central term that is invisible to vector fields
on spacetime.

\subsection{Rotation on Vectors}

For \(J_{ab}=x_a\partial_b-x_b\partial_a\), compute against a translation:
\[
\begin{aligned}
  [J_{ab},P_c]
  &=
    -\partial_c(x_a)\partial_b
    +\partial_c(x_b)\partial_a\\
  &=
    \delta_{bc}P_a-\delta_{ac}P_b.
\end{aligned}
\]
The same calculation applies to boosts because \(J_{ab}\) does not act on \(t\):
\[
  [J_{ab},G_c]
  =
  \delta_{bc}G_a-\delta_{ac}G_b.
\]
Thus both \(P_c\) and \(G_c\) transform as spatial vectors.

\subsection{Rotation-Rotation Bracket}

For two rotations, expanding the vector-field bracket gives
\[
\begin{aligned}
  [J_{ab},J_{cd}]
  &=
    \delta_{bc}J_{ad}
    -\delta_{ac}J_{bd}\\
  &\quad
    +\delta_{ad}J_{bc}
    -\delta_{bd}J_{ac},
\end{aligned}
\]
with the convention \(J_{ba}=-J_{ab}\) and \(J_{aa}=0\).  In \(d=3\), one
representative checked bracket is
\[
  [J_{12},J_{23}]=J_{13}.
\]

\subsection{Checked Representatives}

The test suite pins the signs with the following representative outcomes:
\[
\begin{array}{rcl}
  [H,G_1]&=&P_1,\\
  {}[G_1,H]&=&-P_1,\\
  {}[J_{12},P_1]&=&-P_2,\\
  {}[J_{12},P_2]&=&P_1,\\
  {}[J_{12},G_1]&=&-G_2,\\
  {}[J_{12},J_{23}]&=&J_{13}.
\end{array}
\]
It also checks that \([P_1,G_2]=0\) and \([G_1,G_2]=0\) in the unextended
vector-field algebra.

\subsection{Boundary}

The signs here are signs for vector fields.  A quantum model uses self-adjoint
generators and unitary flows, so translating this table into operator
commutators requires the Stone convention and the chosen \(i[A,B]\) observable
action.  The report therefore cites this shard for the geometric algebra and
CA-20 for the central quantum coefficient.
