% SHARD-ID: CA-22-GALILEAN-COMPILER-INTERFACE
% SHARD-TITLE: Compiler Rules and Examples for Galilean Targets
% SHARD-SUMMARY: Extends the lattice-symmetry compiler with Galilean-specific inputs: mass density, conserved charge, current witness, and central-extension data.
% SHARD-SUMMARY: Gives small successful and failed examples showing why a Hamiltonian density alone cannot compile Galilean boosts.
% SHARD-SUMMARY: Records the next source and proof obligations before Galilean lattice candidates can be promoted to continuum symmetry claims.
% SHARD-KEYWORDS: Galilean compiler, mass density, current witness, central extension, examples, compile error

\section{Compiler Rules and Examples for Galilean Targets}
\label{sec:galilean-compiler-interface}

\subsection{Status, Sources, and Dependencies}

\textbf{Status: compiler specification shard.}  This shard does not implement
the compiler.  It specifies the extra fields a Galilean target requires, using
the checked algebra and first-moment derivations from CA-19--CA-21.

Dependencies: CA-08, CA-16, and CA-18--CA-21.

% Source: references/cft/Schottenloher2008/Schottenloher2008.md:1443 -- Galilei group for free particles in classical nonrelativistic mechanics.
% Source: references/cft/Schottenloher2008/Schottenloher2008.md:1618 -- central-extension lift for projective representations.
% Source: literature/md/cond-mat_9906453/cond-mat_9906453.md:39 -- Galilean-invariant interacting fermions of mass m.
% Source: literature/md/cond-mat_9906453/cond-mat_9906453.md:41 -- center-of-mass interpretation.
% Source: literature/md/2010.11121/2010.11121.md:228 -- continuum time-zero net includes strongly continuous spatial/time translation representation.
% Checked: test/runtests.jl, "Galilei vector-field brackets", "Galilei mass central coefficient", and "first-moment continuity current coefficients".

\subsection{Target Declaration}

A Galilean compiler target is a structured request:
\[
  \operatorname{Target}(\mathrm{Gal}(d),\ \mathrm{projective},\ \mathrm{mass}).
\]
It asks for:
\[
\begin{array}{ll}
H & \text{time translation},\\
P_a & \text{spatial translations / momenta},\\
J_{ab} & \text{rotations},\\
G_a & \text{Galilean boosts},\\
M & \text{central mass generator}.
\end{array}
\]
The finite candidate layer is allowed to output formulas and witnesses.  It is
not allowed to assert continuum Galilean covariance without scaling-limit
evidence.

\subsection{Required Input Slots}

Compared with CA-16, the Galilean target adds the following required slots:
\[
\begin{array}{ll}
\mu_x\text{ or }n_x
  & \text{local mass or particle-number density},\\
M=\sum_x\mu_x
  & \text{conserved total mass/number},\\
{}[H,M]=0
  & \text{finite commutator witness},\\
J_{x,e}
  & \text{current witness for a discrete continuity equation},\\
m\text{ or }M
  & \text{central charge/operator for the projective representation},\\
\text{boundary policy}
  & \text{open, periodic with branch, or thermodynamic-limit handling}.
\end{array}
\]
If any of the first four are absent, boosts and momenta for a Galilean target
are underdetermined.

\subsection{Generated Candidates}

For an open chain, the deterministic candidate formulas are
\[
  G^{\mathrm{lat}}(0)=\sum_j x_j\mu_j,
\]
and, when a continuity witness is supplied,
\[
  P^{\mathrm{lat}}
  :=
  i[H,G^{\mathrm{lat}}(0)]
  =
  \sum_{j=1}^{N-1}(x_{j+1}-x_j)J_{j+1/2}
  \quad+\quad\text{boundary terms}.
\]
Rotations in higher dimension require momentum or current-density data:
\[
  J_{ab}^{\mathrm{lat}}
  \approx
  \sum_x(x_a p_b(x)-x_b p_a(x)).
\]
Thus rotations are second-stage candidates: first compile current/momentum
density, then form the spatial first moment.

\subsection{Example 1: One Free Particle}

Input:
\[
  \mathcal H=L^2(\mathbb R^d),\qquad
  H=\frac{P^2}{2m},\qquad
  [X_a,P_b]=i\delta_{ab}I.
\]
Output at \(t=0\):
\[
  G_a=mX_a,\qquad
  [G_a,P_b]=im\delta_{ab}I,\qquad
  i[H,G_a]=P_a.
\]
Status: checked formal derivation on a common core, with unbounded-operator
domain work deferred.

\subsection{Example 2: Number-Conserving Open Chain}

Input:
\[
  H,\quad n_j,\quad [H,\sum_j n_j]=0,\quad
  i[H,n_j]=J_{j-1/2}-J_{j+1/2}.
\]
For identical mass \(m\),
\[
  \mu_j=mn_j,\qquad
  G^{\mathrm{lat}}(0)=m\sum_j x_jn_j.
\]
The compiler may output the current-sum momentum candidate with coefficient
\(m(x_{j+1}-x_j)\).  It must still mark continuum Galilean covariance as a proof
obligation.

\subsection{Example 3: Hamiltonian Density Only}

Input:
\[
  H=\sum_j h_j
\]
with no \(n_j\), no \(\mu_j\), and no current witness.  For a Lorentzian target,
CA-12 permits the energy-moment proposal.  For a Galilean target, the compiler
must return:
\[
  \texttt{CompileError: missing mass/number density and continuity witness.}
\]
Using \(\sum_jx_jh_j\) here would be a convention bug.

\subsection{Example 4: Spin Chain with No Particle Number}

A spin chain may have a perfectly good Hamiltonian density and exact lattice
translations.  Unless a conserved density is explicitly declared as mass or
particle number, it does not compile as a Galilean particle system.  The correct
output is either a failed Galilean target or a different target declaration.

\subsection{Example 5: Direct Sum of Mass Sectors}

For
\[
  \mathcal H=\mathcal H_{m_1}\oplus\mathcal H_{m_2},
\]
the central generator is not one scalar if \(m_1\ne m_2\).  It is
\[
  M=m_1I_{\mathcal H_{m_1}}\oplus m_2I_{\mathcal H_{m_2}}.
\]
The compiler can carry the OR carrier, but an irreducible fixed-mass sector
selection is a separate projection/sector step in the sense of CA-06.

\subsection{Next Obligations}

The next Galilean work needs:
\[
\begin{array}{ll}
1.&\text{a registered source for the standard mass central extension},\\
2.&\text{model examples with explicit lattice number-current continuity},\\
3.&\text{finite commutator tests for those models},\\
4.&\text{scaling maps proving that current sums converge to continuum momenta},\\
5.&\text{a decision on periodic boundary branch conventions.}
\end{array}
\]
Only after these are present should the compiler promote Galilean lattice
candidates from formulas to continuum symmetry evidence.
