% SHARD-ID: CA-23-GAUSSIAN-BOSON-LORENTZ-ROADMAP
% SHARD-TITLE: Gaussian Boson Lorentz Generator Roadmap
% SHARD-SUMMARY: Opens the Gaussian-boson Lorentz block for spatial dimensions d=1,2,3 and fixes the first solvable target class.
% SHARD-SUMMARY: Decomposes the multi-shard quest into symbol calculus, diagonalisation, generator algebra, Lorentz conditions, numerical checks, and continuum proof obligations.
% SHARD-SUMMARY: States the end goal: identify when translation-invariant Gaussian lattice coefficients scale to the Klein-Gordon/free scalar Poincare representation.
% SHARD-KEYWORDS: Gaussian boson, Lorentz symmetry, Klein-Gordon, lattice generators, roadmap, quadratic Hamiltonian

\section{Gaussian Boson Lorentz Generator Roadmap}
\label{sec:gaussian-boson-lorentz-roadmap}

\subsection{Status, Sources, and Dependencies}

\textbf{Status: roadmap shard with checked seed.}  This shard scopes a
multi-shard programme.  CA-24 supplies the first checked symbol calculation.
The stronger continuum conclusion is not yet proved for a general coefficient
class.

Dependencies: CA-10--CA-17 for lattice Lorentz motivation and acceptance tests,
and CONVENTIONS.md (j).  This shard starts the Gaussian-boson block.

% Source: literature/md/2010.11121/2010.11121.md:13 -- scaling limit of free lattice ground states extends to the massive continuum free field with spacetime translations.
% Source: literature/md/2010.11121/2010.11121.md:51 -- harmonic free lattice fields on the torus.
% Source: literature/md/2010.11121/2010.11121.md:57 -- renormalization condition yields continuum free field of mass m.
% Source: literature/md/2010.11121/2010.11121.md:61 -- construction gives local net of continuum free field in arbitrary dimensions.
% Source: literature/md/2010.11121/2010.11121.md:164 -- free scalar field can be treated with full control.
% Source: literature/md/2010.11121/2010.11121.md:274--293 -- finite torus lattice, dual lattice, and Fourier transform normalization.
% Source: references/cft/Schottenloher2008/Schottenloher2008.md:4186 -- free bosonic QFT via Klein-Gordon operator, arbitrary D>=2.
% Source: references/cft/Schottenloher2008/Schottenloher2008.md:4244 -- Poincare representation for the free boson.

\subsection{Problem Statement}

We now restrict the lattice-symmetry problem to a solvable class:
translation-invariant Gaussian bosonic lattice systems in spatial dimensions
\[
  d=1,2,3,
  \qquad D=d+1.
\]
The physical target is Lorentz/Poincare covariance in the scaling limit.  The
expected endpoint is the scalar free field:
\[
  \omega(k)^2=m^2+|k|^2
\]
after the correct normalization of lattice spacing, velocity, field strength,
and mass.

The inputs are not merely a matrix Hamiltonian.  The compiler receives:
\[
\begin{array}{ll}
\Lambda_{\epsilon,L} & \text{periodic hypercubic lattice or infinite lattice},\\
q_x,p_x & \text{canonical bosonic field variables},\\
V_r & \text{translation-invariant quasi-local quadratic coefficients},\\
\epsilon & \text{lattice spacing},\\
S_\epsilon & \text{state/scaling-map data, when a continuum claim is requested}.
\end{array}
\]
The first coefficient tier is the scalar canonical Hamiltonian
\[
  H=\frac{1}{2}\sum_xp_x^2+\frac{1}{2}\sum_{x,r}q_xV_rq_{x+r}.
\]
Full bosonic BdG/pairing systems are a later tier, not silently included in the
first convention.

\subsection{North Star}

For each admissible coefficient family \(V^{(\epsilon)}_r\), compute the
candidate generators as functions of the coefficients:
\[
  H,\quad P_a,\quad J_{ab},\quad K_a,
\]
then compute the residuals in the Poincare algebra.  The residuals become
conditions on the symbol
\[
  \omega_\epsilon(k)^2=\sum_rV^{(\epsilon)}_r e^{i\epsilon k\cdot r}.
\]
The end goal is a theorem-shaped statement:

\[
\begin{gathered}
\text{Gaussian lattice coefficients satisfying locality, positivity,}\\
\text{single low-energy minimum, and isotropic Klein-Gordon Taylor data}
\\
\Longrightarrow
\text{ scaling limit is the free scalar/Klein-Gordon Poincare system.}
\end{gathered}
\]

\textbf{CA-15 proof boundary.}  This is a theorem-shaped target, not a theorem
already obtained from the symbol checks.  A continuum claim still needs the
CA-15 payload: scaling maps, state sequence, topology, operator domains, local
observable covariance, and boundary/renormalization policy.  The value of the
Gaussian block is that the algebraic residuals can be made explicit before
attempting that proof.

\subsection{Shard Roadmap}

% Source: report/README.md:43--49 -- current include order for CA-23--CA-29.
% Source: report/SHARD_CATALOG.md:221--272 -- current CA-24--CA-29 titles and summaries.
% Source: reviews/2026-05-30_gaussian_lorentz/ACTION_PLAN.md:179--194 -- R2 roadmap repair scope.
% Source: reviews/2026-05-30_gaussian_lorentz/ORCHESTRATION.md:38--46 -- queued source and stress-energy follow-up labels.
The current Gaussian block is:
\begin{description}
\item[CA-24.]
Symbol calculus for the scalar coefficient tier, including the boost-time
residual and the checked nearest-neighbour Klein-Gordon seed.
\item[CA-25.]
Diagonalization into one-particle symbols, with the sourced free-scalar Fock
construction separated from the local coefficient-symbol derivation and BdG
pairing deferred.
\item[CA-26.]
Smooth-patch one-particle generator algebra for \(H,P,J,K\), exact brackets,
dispersion-dependent residuals, and only formal \(\mathrm{d}\Gamma\)
bookkeeping on the algebraic finite-particle core.
\item[CA-27.]
Lorentz residual conditions on coefficients: Taylor constraints, anisotropic
failures, and doubler counterexamples.
\item[CA-28.]
Numerical invariant suite for scalar symbols and finite periodic stiffness
matrices, covering Klein-Gordon, anisotropic, and doubler examples.
\item[CA-29.]
Real-space energy-density convention for the scalar Gaussian Hamiltonian:
cell volume, bond sharing, boundary policy, vacuum subtraction, and improvement
terms before any position-weighted use.
\end{description}

The queued follow-up block is proposal work, not completed stress-energy
material:
\begin{description}
\item[1+1 stress candidates.]
Build candidate lattice stress-energy formulas only after the CA-29 density and
a registered current convention are in place.
\item[Current-vs-symbol equivalence.]
Check when real-space first moments and currents reproduce the flat local-symbol
residuals used in CA-24--CA-27.
\item[Higher-dimensional cells.]
Propose cell-current formulas in \(d=2,3\), keeping orientation, boundary, and
improvement choices explicit.
\item[Stress-energy numerical suite.]
Add numerical checks only after candidate formulas and source-backed expected
identities exist.
\item[Source acquisition.]
Register the continuum stress-energy/Poincare-generator source, harmonic-chain
energy-current formulas, and the discrete Gaussian/free-field Virasoro source.
\end{description}

\subsection{Side Quests}

\textbf{Source acquisition.}  The registered local OAR sources cover the free
scalar lattice scaling limit and spacetime translations.  They explicitly warn
that Lorentz/conformal convergence requires further work.  We need a registered
source for the exact continuum stress-energy/Poincare-generator formulas, and a
registered copy of the discrete Gaussian free-field Virasoro source currently
present only as bibliography metadata.

\textbf{Generator ambiguity.}  A local Hamiltonian density is not unique:
\[
  h_x\mapsto h_x+\nabla\cdot b_x
\]
changes first moments by boundary or current terms.  The Gaussian block must
track which density convention produced \(K_a=\sum_xx_ah_x\).

\textbf{Periodic positions.}  Numerical tests use periodic lattices and finite
Brillouin grids.  First moments, position coordinates, and finite-grid
momentum-coordinate generators need a branch, sawtooth, Fourier interpolation,
or another explicit periodic coordinate convention.  Until that is fixed, early
finite tests are symbol/eigenvalue checks on the periodic grid, or open-chain
first-moment checks.

\textbf{BdG pairing.}  A general quasi-local Hamiltonian in creation and
annihilation operators may include pairing terms.  The scalar \(q,p\) tier
captures Klein-Gordon directly; BdG requires symplectic positivity and a stable
bosonic Bogoliubov convention before generator formulas are promoted.

\subsection{Numerical Verification Plan}

The Julia layer should stay at the invariant level:
\[
\begin{array}{ll}
1.&\text{symbol from coefficients equals the analytic dispersion},\\
2.&\frac12\nabla\omega_\epsilon(k)^2\text{ equals the smooth-patch boost-time residual},\\
3.&\frac{\sin(\epsilon k_a)}{\epsilon}\to k_a\text{ for the KG lattice},\\
4.&\text{anisotropic Hessians fail the Lorentz residual checks},\\
5.&\text{finite Fourier matrices diagonalize periodic stiffness matrices}.
\end{array}
\]
The first three checks are implemented with CA-24.

\subsection{Acceptance Boundary}

The roadmap rows above remain scoped by the CA-15 proof boundary stated next to
the theorem-shaped goal.  Passing symbol or finite-matrix tests is not yet a
proof of Lorentz invariance.
