% SHARD-ID: CA-26-GAUSSIAN-BOSON-GENERATOR-ALGEBRA
% SHARD-TITLE: One-Particle Generator Algebra for Gaussian Bosons
% SHARD-SUMMARY: Defines the one-particle H, P, J, and K candidates on a smooth momentum-space core.
% SHARD-SUMMARY: Derives the commutators that are exact for any scalar dispersion and isolates the residuals that depend on omega.
% SHARD-SUMMARY: Records only formal dGamma bookkeeping on an algebraic finite-particle core, leaving many-body representation theorems open.
% SHARD-KEYWORDS: Gaussian boson, Poincare generators, boost, rotation, commutator, second quantization

\section{One-Particle Generator Algebra for Gaussian Bosons}
\label{sec:gaussian-boson-generator-algebra}

\subsection{Status, Sources, and Dependencies}

\textbf{Status: flat local derivation with limited Julia checks.}  The Poincare
target, Stone convention, and free-boson Fock realization are sourced.  The
commutator identities in this shard are local differential-operator derivations
on a flat \(L^2(dk)\) smooth momentum-space core.  The Julia suite currently
checks coefficient/symbol consequences: Klein-Gordon symbols, boost-time
symbols, low-energy Hessian examples, finite periodic massive spectra, the
massless doubler coefficient-rejection witness, and the Gaussian tolerance
policy.  It does not check the full generator bracket table or any
second-quantized identity.  The smooth-core derivation is not yet a sourced OAR
or mass-shell Poincare representation theorem.

Dependencies: CA-11, CA-24--CA-25, and CONVENTIONS.md (h), (j), (k).

% Source: references/cft/Schottenloher2008/Schottenloher2008.md:4040--4050 -- Stone convention for one-parameter unitary groups.
% Source: references/cft/Schottenloher2008/Schottenloher2008.md:4092 -- Poincare group and relativistic invariance.
% Source: references/cft/Schottenloher2008/Schottenloher2008.md:4106 -- unitary Poincare representation.
% Source: references/cft/Schottenloher2008/Schottenloher2008.md:4113--4117 -- translation generators P_mu and P_0=H.
% Source: references/cft/Schottenloher2008/Schottenloher2008.md:4244 -- free bosonic QFT carries a natural Poincare action on Fock space.
% Source: literature/md/2010.11121/2010.11121.md:274--293 -- finite torus lattice, dual lattice, and Fourier transform normalization.
% Source: literature/md/2010.11121/2010.11121.md:648--681 -- OAR scalar product, flat Fock realization, and implemented Klein-Gordon dynamics.
% Convention: CONVENTIONS.md (j) -- flat L^2(dk) status of the CA-24--CA-26 local symbol algebra.
% Convention: CONVENTIONS.md (k) -- vector-field to self-adjoint-generator sign map.
% Checked: test/runtests.jl, "Gaussian boson numerical tolerance policy",
%   "Gaussian boson Klein-Gordon symbols", "Gaussian boson boost-time symbols",
%   "Gaussian boson Lorentz Hessian examples",
%   "Gaussian boson finite periodic massive coefficient spectra", and
%   "Gaussian boson massless doubler coefficient rejection".

\subsection{Momentum Core}

There are two separate settings.

\textbf{Smooth patch.}  Work on a connected smooth momentum patch \(\mathcal U\) where
\(\omega(k)>0\), with the flat measure \(dk\).  Let
\[
  \mathcal D_0=C_c^\infty(\mathcal U)\subset L^2(\mathcal U,dk)
\]
and define
\[
  X_a=i\partial_{k_a},
  \qquad
  h=\omega(k),
  \qquad
  p_a=k_a .
\]
The one-particle Hamiltonian and momenta are multiplication operators:
\[
  H^{(1)}=h,\qquad P_a^{(1)}=p_a.
\]
The rotation and time-zero boost candidates are
\[
  J_{ab}^{(1)}=X_aP_b-X_bP_a,
  \qquad
  K_a^{(1)}=\frac{1}{2}(X_ah+hX_a).
\]
These are written in the self-adjoint sign convention of CONVENTIONS.md (k);
in particular \(J_{ab}\) is the Gaussian rotation sign, not the raw lowered-index
CA-11 vector field \(M_{ab}\).
All commutators below are identities on \(\mathcal D_0\); promoting them to
self-adjoint generator statements is a separate domain theorem.  Promoting this
flat formula to the OAR one-particle Hilbert space or Schottenloher's
mass-shell/free-boson Poincare representation additionally requires a sourced
or locally derived unitary intertwiner.

\textbf{Finite periodic grids.}  On a finite periodic Gaussian lattice, the
dual grid is used for multiplication symbols and Fourier eigenvalue checks.
The current finite suite does not place the differential operator
\(X_a=i\partial_{k_a}\) on that grid and does not claim an exact finite
\([X_a,p_b]=i\delta_{ab}\) commutator.  Periodic first-moment or
momentum-coordinate generator tests require an explicit branch, sawtooth,
Fourier-interpolation, or comparable periodic-coordinate convention before
they can be stated.  Thus the exact differential commutators below are
smooth-patch algebra; the finite periodic checks remain symbol/eigenvalue
checks.

\subsection{Exact Translation and Rotation Brackets}

On the smooth core, because \(h\) and \(p_a\) are multiplication operators,
\[
  i[h,p_a]=0.
\]
The canonical differential identity is the smooth-patch identity
\[
  [X_a,p_b]=i\delta_{ab}.
\]
Therefore
\[
  i[P_c,J_{ab}]
  =
  \delta_{ac}P_b-\delta_{bc}P_a.
\]
This is independent of the dispersion.  It is kinematical on the smooth patch:
it comes from the chosen momentum coordinate and the differential
representation of rotations, in the rotation sign convention of
CONVENTIONS.md (k).

\subsection{Boost-Momentum Bracket}

For the boost candidate,
\[
\begin{aligned}
  [P_b,K_a]
  &=
  \frac{1}{2}\bigl([P_b,X_a]h+h[P_b,X_a]\bigr)\\
  &=
  -i\delta_{ab}h.
\end{aligned}
\]
Hence
\[
  i[P_b,K_a]=\delta_{ab}H .
\]
This is the boost-momentum line of the self-adjoint target table in
CONVENTIONS.md (k).
On the smooth patch, this part of the Poincare algebra is exact for every
positive scalar dispersion because it only uses \(K_a\) as the symmetrized first
moment of energy.

\subsection{Boost-Time Residual}

CA-24 derived
\[
  i[H,K_a]
  =
  \frac{1}{2}\partial_a\omega(k)^2.
\]
The continuum target with speed \(c=1\) wants \(P_a=k_a\).  The residual is
\[
  R_a(k)
  :=
  \frac{1}{2}\partial_a\omega(k)^2-k_a.
\]
For speed \(c\), replace \(k_a\) by \(c^2k_a\).  This is the first algebraic
filter on Hamiltonian coefficients, and it is the only generator-algebra
residual currently checked at coefficient-symbol level in Julia.

\subsection{Rotation-Hamiltonian Residual}

The Hamiltonian commutes with rotations iff the dispersion is radial on the
patch:
\[
  i[h,J_{ab}]
  =
  k_b\partial_a\omega-k_a\partial_b\omega.
\]
Equivalently,
\[
  k_b\partial_a\omega^2-k_a\partial_b\omega^2=0
\]
where \(\omega>0\).  Hypercubic lattice symbols usually fail this globally on
the Brillouin zone but may satisfy it to leading order at a low-energy minimum.

\subsection{Boost-Boost Residual}

Write
\[
  B_a(k):=\frac{1}{2}\partial_a\omega(k)^2,
  \qquad
  R_a(k):=B_a(k)-k_a.
\]
Equivalently \(B_a=h\partial_a h\).  Since
\[
  K_a=i\left(h\partial_a+\frac{1}{2}\partial_ah\right),
\]
the commutator of the first-order differential parts gives
\[
  i[K_a,K_b]
  =
  J_{ab}+R_bX_a-R_aX_b.
\]
The leading \(+J_{ab}\) is the sign dictated by CONVENTIONS.md (k): CA-11 has
\([M_{0a},M_{0b}]=-M_{ab}\), while the self-adjoint rotation here is
\(J_{ab}=-A_{M_{ab}}\).
Here the possible divergence term vanishes because \(R\) is a gradient
residual:
\(\partial_aR_b=\partial_bR_a\).  Thus the boost-boost bracket closes to the
rotation bracket whenever the boost-time residual \(R_a\) vanishes on the patch.
This boost-boost statement is a smooth-core algebraic derivation only; no Julia
test currently checks the boost-boost operator bracket.  The remaining caveat is
analytic, not algebraic: all of these are unbounded operators and still require a
common self-adjointness/core theorem before this becomes a continuum
representation statement.

\subsection{Second Quantization}

Let
\[
  \mathcal F_{\mathrm{alg}}(\mathcal D_0)
  :=
  \bigoplus_{n=0}^{\mathrm{fin}}\mathcal D_0^{\otimes_s n}
  \subset \mathcal F_+\bigl(L^2(\mathcal U,dk)\bigr)
\]
be the algebraic finite-particle core.  The formal many-body candidates are
\[
  H=d\Gamma(h),\quad
  P_a=d\Gamma(p_a),\quad
  J_{ab}=d\Gamma(J_{ab}^{(1)}),\quad
  K_a=d\Gamma(K_a^{(1)}),
\]
with a separate vacuum-energy convention if absolute Hamiltonian values are
used.  The formal core identity is
\[
  [d\Gamma(A),d\Gamma(B)]\psi=d\Gamma([A,B])\psi,
  \qquad \psi\in\mathcal F_{\mathrm{alg}}(\mathcal D_0),
\]
whenever the one-particle commutator identity is already established on
\(\mathcal D_0\) and the displayed products preserve that algebraic domain.
This is a local \(d\Gamma\) bookkeeping statement on the algebraic
finite-particle core, not a Julia-checked invariant and not a self-adjoint
many-body representation theorem.  The current tests do not lift residuals to
Gaussian Fock space; domain/self-adjointness for the unbounded \(d\Gamma\)
generators and the sourced OAR/mass-shell bridge remain open proof obligations.
